Aneuploidy can be an evolutionary detour \\ on the path to adaptation

### 2022-09-19 ms/ms.tex
small revisions with AY

% Title page
% Chromosomal duplication is a transient evolutionary solution to stress
\title{Chromosomal duplication can be an evolutionary detour on the path to adaptation}

@@

Keywords: aneuploidy, evolutionary model, adaptive evolution

@@

Together, our results suggest that aneuploidy can be an evolutionary "detour" rather than a "stepping stone": it can delay, rather than facilitate, the adaptation of the population, and cells that become aneuploid may leave less descendants compared to cells that remain diploid.
\end{abstract}

@@

% Conclusions
\paragraph{Conclusions.}
Here, we tested the hypothesis that aneuploid cells are an evolutionary "stepping stone", or adaptive intermediate, between wild-type euploid cells and mutant euploid cells. 
Our results suggest that, although it seems the population goes from euploid to aneuploid and back, this is not the case at the individual level. We estimate that only about 10-15\% of the euploid cells descended from aneuploid cells, whereas the rest are direct descendants of the wild-type euploid cells.
This surprising result reinforces the importance of models when making interpretations on evolutionary processes, and emphasizes the unintuitive outcomes of clonal interference during adaptive evolution.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{DNA sequencing.}
Whole-genome sequencing of the ancestral diploid strain (\euwt) was performed on a sample generated from a single colony of the ancestor.
Whole-genome sequencing of the four evolving populations ($H2$ after 450 and 2,350 generations, and $H4$ after 450 and 1,700 generations) was performed at the population level.
A liquid sample from each evolving population was taken for sequencing from each culture as it grows under the same conditions as the lab evolution experiments (YPD at \SI{39}{\celsius}).
DNA was purified from the collected cells for each population and downstream deep sequencing was performed according to the Illumina protocol for amplicon sequencing.

>>>>

% Title page
% Chromosomal duplication is a transient evolutionary solution to stress
% \title{Chromosomal duplication can be an evolutionary detour on the path to adaptation}
\title{Aneuploidy can be an evolutionary detour \\ on the path to adaptation}

@@

Keywords: whole-chromosome duplication, evolutionary model, adaptive evolution

@@

Together, our results suggest that aneuploidy can be an inevitable evolutionary "detour" rather than a "stepping stone": it can delay, rather than facilitate, the adaptation of the population, and cells that become aneuploid may leave less descendants compared to cells that remain diploid.
\end{abstract}

@@

% Conclusions
\paragraph{Conclusions.}
Here, we tested the hypothesis that aneuploid cells are an evolutionary "stepping stone", or adaptive intermediate, between wild-type euploid cells and mutant euploid cells. 
Our results suggest that, although it seems the population goes from euploid to aneuploid and back, this is not the case at the individual level. We estimate that only about 10-15\% of the euploid cells descended from aneuploid cells, whereas the rest are direct descendants of the wild-type euploid cells. Thus, aneuploidy can delay, rather than accelerate, adaptation, and and cells that become aneuploid may leave less descendants than cells that remain euploid. 
This surprising result reinforces the importance of models when making interpretations on evolutionary processes, and emphasizes the unintuitive outcomes of clonal interference during adaptive evolution.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{DNA sequencing.}
Whole-genome sequencing of the ancestral diploid strain (\euwt) was performed on a sample from a single colony of the ancestor.
Whole-genome sequencing of the four evolving populations ($H2$ after 450 and 2,350 generations, and $H4$ after 450 and 1,700 generations) was performed on a sample from these populations (rather than from single colonies).
The liquid sample from each evolving population was taken from each culture as it grows under the same conditions as the lab evolution experiments (YPD at \SI{39}{\celsius}).
DNA was purified from the collected cells for each population and downstream deep sequencing was performed according to the Illumina protocol for amplicon sequencing.
